Stochastic Variance-Reduced Policy Gradient

### 2018-04-16 rebuttals.txt

We would like to thank all the reviewers for their feedback and in-depth comments.

Reviewer 1:

The heuristics proposed in the paper were not developed with the main purpose of beating the baseline, but to make the algorithm work (learn) in a stable way. The intrinsic difficulties of adapting SVRG to the RL framework have, as a consequence, a certain sensitivity to the meta-parameters. Our contribution should be intended as two-fold: a theoretical part supporting the use of SVRG in RL, and a more empirical part highlighting the additional challenges of variance reduction in RL (where we also show how these difficulties can be overcome with proper heuristics). We think the community would benefit from a picture of both the opportunities and the risks of applying SVRG to policy gradient learning.

@@

The proof to Lemma 3.1 was meant to be in the appendix and is missing by mistake. We will definitively include it. The variance property is true at convergence because the O(1/B) terms cancel out as both the parameter and the snapshot parameter approach the optimum. More formally: "As \theta -> \theta^*, also \tilde\theta -> \theta^*; hence, by continuity of J: Var(SVRG) -> Var(\hat\nablaJ) + 1/B*Var[g(\theta^*) - 1*g(\theta^*)] = Var(\hat\nabla)". The importance weight is one because the two parameters are the same. See also a similar proof in Section 2 of the original SVRG paper.

The performance of GPOMDP on Swimmer is lower than the one reported in Duan et al. because we do not employ the linear critic used in that paper. Since critic and SVRG are both ways to reduce the variance of the gradient estimate, and a combination of the two is non-trivial (how do you critique snapshot policies?) we turned the critic off in both GPOMDP and SVRPG to measure the net advantage of SVRG. Later, in the Half-Cheetah experiment, we include the critic in both algorithms.

We reported Duan et al. values only for the Half-Cheetah experiment because it is the only case in which we used the critic. This represents a further step: evaluating the combination of SVRG and critic as variance reduction techniques. We recognize that the current presentation is confusing. In fact, we now have reproduced the GPOMDP/Half-Cheetah result and we will simply report our results in the final version. We will mention that experiments on Half-Cheetah are coherent with Duan et al., while the ones on Cart-Pole and Swimmer cannot be compared since we did not use the critic.

@@

The version of Cart-Pole we use is the same of Duan et al, and this should be mentioned, although the task is thoroughly described in Appendix E. As for Swimmer, the lower scores are due to the fact that we do not include a critic in either GPOMDP or SVRPG.

As a first attempt to bring the SVRG technique to the policy optimization framework, we do not claim SVRPG has the maturity of algorithms like TRPO. Recent policy optimization algorithms employ a series of tools, e.g. natural gradients and exploration bonuses, that are outside the scope of this paper but represent promising future-work material.

We thank the reviewer for pointing out typos.

Reviewer 2:
We see that meta-parameters are a delicate aspect of our algorithm. The pragmatic role of Theorem 4 is to show the interplay of these parameters more than prescript specific values. In our work, we both provide general theoretical guarantees and propose a practical way to select meta-parameters. The statement of Theorem 4 (and some subtleties of its proof) is heavily guided the choice of this heuristics, and we will make this more clear in the final version. More specifically:

1) Variance reduction algorithms for supervised learning often provide explicit guidelines on how to set the epoch size, e.g. O(N). In RL, this meta-parameter is more delicate since it is linked to the variance introduced by importance sampling. The theory would prescribe a large epoch size to avoid instabilities, but this would hinder the efficiency of the algorithm. The value optimizing this tradeoff is difficult to find and changes over time. It is also problem dependent since it also depends on the geometry of policy space. For these reasons, we designed an adaptive algorithm, which is heuristic but based on insights from the theory. In the final version, we will explain this process better and we will show examples of manual selection of m for sensitivity analysis. Future work will try to bridge the gap between theory and practice, e.g. by employing natural gradient or more sophisticated importance weighting techniques to reduce the sensitivity of SVRPG to this meta-parameter. 
2) In nonconvex optimization, it is often the case that the step sizes prescribed by the theory are prohibitively conservative to be used in applications. For instance, in Reddi et al. the authors understand from their theoretical analysis that a constant step-size should be used, but they empirically search for the specific value that yields the best practical performance. Even in the original SVRG paper, common annealing schemas are employed in the experiments.
3) In Theorem 4.1, \theta^* denotes a global optimizer, although only convergence to a local optimum is guaranteed. We will make this more clear.
4) \theta_m^0, \tilde\theta^0 and \theta_0^1 are all the same. Superscripts denote epochs and subscripts inner iterations, as should be clear from the pseudocode. We will check for notation inconsistencies in the text.
5) Also under these assumptions, plain PG is an instance of Stochastic Gradient Ascent, which has a 1/\sqrt{T} rate.
6) We consider the horizon H to be part of the problem, but we should mention that problems with long horizons can be more difficult since we employ importance weighting. In the final version, we will show an example where we vary the horizon of a task for sensitivity analysis.

On experiments:
1) The comparison is fair, figures have the number of trajectories on the horizontal axis, and the extra trajectories collected by  SVRPG at the snapshot are indeed counted. Our algorithm updates the policy less often but manages to be faster anyway because the quality of the updates is higher.
2) Using fixed batch sizes is indeed suboptimal, but we leave this aspect for future work since adapting batch sizes in the policy gradient framework is a  difficult problem that has been only slightly explored in the RL literature so far (cfr. Papini et al., "Adaptive batch size for safe policy gradients." NIPS 2017).
3) This is because we used the rllab variant of Cart-Pole (Duan et al., 2016), where the agent collects a reward of ~10 at each time-step, as described in Appendix E.
4) For SVRPG, \alpha is selected with the Adam variant described in Section 5.2. For GPOMDP, we use Adam. All meta-parameters are reported in Appendix E and have been chosen to make the comparison fair.

### 2018-05-21 svrpg.tex
minor changes/additions suggested by reviewers

%Paper structure
After providing background on policy gradient and SVRG in Section \ref{sec:pre}, in Section \ref{sec:alg} propose SVRPG, a variant of SVRG for the policy gradient framework, addressing all the difficulties mentioned above. In Section \ref{sec:conv} we provide convergence guarantees for our algorithm, and we show a convergence rate that has an $O(\nicefrac{1}{T})$ dependence on the number $T$ of iterations. In Section \ref{sec:stopping} we suggest how to set the meta-parameters of SVRPG, while in Section \ref{sec:prac} we discuss some practical variants of the algorithm. Finally, in Section \ref{sec:exp} we empirically evaluate the performance of our method on popular continuous RL tasks.

@@

\vspace{-0.05in}
\subsection{Stochastic Variance-Reduced Gradient}\label{sec:svrg}
\vspace{-0.05in}
Finite-sum optimization is the problem of maximizing an objective function $f(\vtheta)$ which can be decomposed into the sum or average of a finite number of functions $z_i(\cdot|\vtheta)$:
\begin{align*}
        \max_{\vtheta} \left\{ f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}z_i(\vtheta)\right\}.
\end{align*}
This kind of optimization is very common in machine learning, where each $z_i$ may correspond to a data sample $x_i$ from a dataset $\mathcal{D}_N$ of size $N$ (\ie $z_i(\vtheta) = z(x_i|\vtheta)$). 
%% In this case, we adopt the following, more meaningful notation:
%% \begin{align*}
%% \max_{\vtheta} f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}g(x_i \vert \vtheta).
%% \end{align*}
A common requirement is that $z$ must be smooth and concave in $\vtheta$.\footnote{Note that we are considering a maximization problem instead of the classical minimization one.} 
Under this hypothesis, full gradient (FG) ascent~\citep{cauchy1847methode} with a constant step size achieves a linear convergence rate in the number $T$ of iterations (\ie parameter updates)~\citep{nesterov2013introductory}.
However, each iteration requires $N$ gradient computations, which can be too expensive for large values of $N$. Stochastic Gradient (SG) ascent~\citep[\eg][]{robbins1951stochastic,bottou2004large} overcomes this problem by sampling a single sample $x_i$ per iteration, but a vanishing step size is required to control the variance introduced by sampling. As a consequence, the lower per-iteration cost is paid with a worse, sub-linear convergence rate \cite{nemirovskii1983problem}.
Starting from SAG, a series of variations to SG have been proposed to achieve a better trade-off between convergence speed and cost per iteration: \eg SAG~\citep{roux2012stochastic}, SVRG~\cite{johnson2013accelerating}, SAGA~\cite{defazio2014saga}, and Finito~\cite{defazio2014finito}. 
The common idea is to reuse past gradient computations to reduce the variance of the current estimate.
In particular, Stochastic Variance-Reduced Gradient (SVRG) is often preferred to other similar methods for its limited storage requirements, which is a significant advantage when deep and/or wide neural networks are employed.

@@

\vspace{-0.05in}
\section{SVRG in Reinforcement Learning}\label{sec:alg}
\vspace{-0.05in}
In online RL problems, the usual approach is to tune the batch size of SG to find the optimal trade-off between variance and speed.
Recall that, compared to SL, the samples are not fixed in advance but we need to collect them at each policy change.
Since this operation may be costly, we would like to minimize the number of interactions with the environment.
For these reasons, we would like to apply SVRG to RL problems in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence.
However, a direct application of SVRG to RL is not possible due to the following issues:
% In on-line policy gradient problems, SG is not really a choice, but is dictated by the necessity of interacting with an unknown environment.\todopir{I do not understand the meaning of this sentence. I would say that SG is not a choice by referring to papers that use batches}
% We would like to apply SVRG to the policy gradient framework in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence. In this scenario, the function $f(\vtheta)$ to optimize is $J(\vtheta)$ and the (implicit) dataset $\mathcal{D}$ is the set of all possible trajectories $\mathcal{T}$. Compared to the typical finite sum optimization scenario, we have the following additional challenges:
\begin{description}
        \item[Non-concavity:] the objective function $J(\vtheta)$ is typically non-concave.
        \item[Infinite dataset:] the RL optimization cannot be expressed as a finite-sum problem. The objective function is an expected value over the trajectory density $p_{\vtheta}(\tau)$ of the total discounted reward, for which we would need an infinite dataset.
%    \item[Approximation:] the dataset $\mathcal{T}$ can be infinite, but we can only sample a finite number of trajectories;
        \item[Non-stationarity:] the distribution of the samples changes over time. In particular, the value of the policy parameter $\vtheta$ influences the sampling process.
\end{description}
To deal with non-concavity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as Gaussian\footnote{See Appendix~\ref{app:gauss} for more details on the Gaussian policy case.} and softmax~\citep[\eg][]{Furmston2012unifying,pirotta2015lipschitz}.
Because of the infinite dataset, we can only rely on an estimate of the full gradient.
\citet{harikandeh2015stopwasting} analysed this scenario under the assumptions of $z$ being concave, showing that SVRG is robust to an inexact computation of the full gradient. In particular, it is still possible to recover the original convergence rate if the error decreases at an appropriate rate. In Section \ref{sec:conv} we will show how the estimation accuracy impacts on the convergence results with a non-concave objective.
Finally, the non-stationarity of the optimization problem introduces a bias into the SVRG estimator in Eq.~\eqref{E:svrg.gradient.correction}.
To overcome this limitation we employ importance weighting~\citep[\eg][]{rubinstein1981simulation,precup2000eligibility} to correct the distribution shift.
%Finally, we employ importance weighting~\citep[\eg]{rubinstein1981simulation,precup2000eligibility} to guarantee proper sampling of trajectories.

\begin{algorithm}[tb]
	\caption{SVRPG}
	\label{alg:svrpg}
	\begin{algorithmic}
		\STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, batch size $N$, mini-batch size $B$, gradient estimator $g$, initial parameter $\vtheta_{m}^0 := \wt{\vtheta}^0$
		\FOR{$s=0$ {\bfseries to} $S-1$}
		\STATE $\vtheta_0^{s+1} := \wt{\vtheta}^{s} = \vtheta_{m}^s$
		\STATE Sample $N$ trajectories $\{\tau_j\}$ from $p(\cdot\vert\wt{\vtheta}^{s})$
		\STATE $ \wt{\mu} = \gradApp{\wt{\vtheta}^{s}}{N}$ (see Eq.~\eqref{E:policygradient.estimate})% = \frac{1}{N}\sum_{j=0}^{N-1}g(\tau_j | \wt{\vtheta}^s)$
		\FOR{$t=0$ {\bfseries to} $m-1$}
		\STATE Sample $B$ trajectories $\{\tau_i\}$ from $p(\cdot\vert\vtheta_t^{s+1})$
		%        \STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1} \left( g(\tau_i|\vtheta_t^{s+1}) - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \right)$
		\STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1}
		\begin{aligned}[t]
		\Big( & g(\tau_i|\vtheta_t^{s+1})\\ 
		& \quad{} - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \Big)
		\end{aligned}$
		%        \STATE $\begin{aligned}[t]
		%                &v^{s+1}_t = \wt{\mu}\\ 
		%                & +\frac{1}{B} \sum\limits_{i=0}^{B-1} \left( g(\tau_i|\vtheta_t^{s+1}) - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \right)
		%                \end{aligned}$
		\STATE $v^{s+1}_t = \wt{\mu} + c^{s+1}_t$ % \gradApp{\wt{\vtheta}^{s}}{N} + c^{s}_t$
		%%		\begin{align*}
		%%		\blacktriangledown J(\vtheta_t^{s+1}) = 
		%%		&\gradApp{\wt{\vtheta}^s}{N} \\
		%%		&+\frac{1}{B}\sum_{i=0}^{B-1}\left[ 
		%%		\score{}{\tau_i\vert\vtheta_t^{s+1}}\Reward(\tau_i)\right. \\
		%%		&\left. - \omega(\tau_i)\score{}{\tau_i \vert \wt{\vtheta}^{s}}\Reward(\tau_i)\right]
		%%		\end{align*}
		\STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v^{s+1}_t$
		\ENDFOR
		\ENDFOR
		\STATE {\bfseries return} $\vtheta_A\coloneqq\vtheta_t^{s+1}$ with $(s,t)$ picked uniformly at random from $\{[0,S-1]\times[0,m-1]\}$
	\end{algorithmic}
\end{algorithm}

@@

\textit{3) Non-stationarity.} 
Similarly to what done in SL~\citep{cortes2010learning}, we require the variance of the importance weight to be bounded.
	\begin{restatable}[On the variance of importance weights]{assumption}{varweights}\label{ass:M2}
		There is a constant $W<\infty$ such that, for each pair of policies encountered in Algorithm~\ref{alg:svrpg} and for each trajectory,
		\[
                \mathbb{V}ar\left[\omega(\tau| \vtheta_1, \vtheta_2)\right] \leq \VARIS, \quad \forall \vtheta_1,\vtheta_2 \in \realspace^d , \tau \sim p(\cdot|\vtheta_1).
		\]
	\end{restatable}
Differently from Assumptions~\ref{ass:bounded_score} and ~\ref{ass:REINFORCE}, Assumption~\ref{ass:M2} must be enforced by a proper handling of the epoch size $m$.

We can now state the convergence guarantees for SVRPG.
\begin{restatable}[Convergence of the SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
Assume the REINFORCE or the G(PO)MDP gradient estimator is used in SVRPG (see Equation~\eqref{E:svrpg.estimate.batch}).
Under Assumptions~\ref{ass:bounded_score}, \ref{ass:REINFORCE} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Algorithm~\ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,\zeta, \xi$ and for proper choice of the step size $\alpha$ and the epoch size $m$, the following property:
\begin{align*}
	&\EVV[]
	{\norm[2]{\nabla J(\vtheta_A)}^2} 
		\leq
		\frac{J(\vtheta^*)-J(\vtheta_0)}{\psi T} +
		\frac{\zeta}{N}
		+\frac{\xi}{B},
\end{align*}
where $\psi,\zeta,\xi$ depend only on $\GRADLOG,\HESSLOG,\VARRF,\VARIS,\alpha$ and $m$.
\end{restatable}
% Specific constants and meta-parameter constraints are provided in the proof of the theorem in Appendix \ref{app:proofs}.
Refer to Appendix~\ref{app:proofs} for a detailed proof involving the definition of the constants and the meta-parameter constraints.
%The theoretical settings may be too conservative.
By analysing the upper-bound in Theorem~\ref{theo:convergence} we observe that: I) the $\mathcal{O}(\nicefrac{1}{T})$ term is coherent with results on non-concave SVRG~\citep[\eg][]{reddi2016stochastic}; II) the $\mathcal{O}(\nicefrac{1}{N})$ term is due to the FG approximation and is analogous to the one in~\citep{harikandeh2015stopwasting}; III) the $\mathcal{O}(\nicefrac{1}{B})$ term is due to importance weighting.
To achieve asymptotic convergence, the batch size $N$ and the mini-batch size $B$ should increase over time. 
In practice, it is enough to choose $N$ and $B$ large enough to make the second and the third term negligible, \ie to mitigate the variance introduced by FG approximation and importance sampling, respectively.
Once the last two terms can be neglected, the number of trajectories needed to achieve $\norm[2]{\gradJ{\vtheta}}^2\leq\epsilon$ is $O(\frac{B+\nicefrac{N}{m}}{\epsilon})$. In this sense, an advantage over batch gradient ascent can be achieved with properly selected meta-parameters. In Section~\ref{sec:stopping} we propose a joint selection of step size $\alpha$ and epoch size $m$.
Finally, from the return statement of Algorithm \ref{alg:svrpg}, it is worth noting that $J(\vtheta_A)$ can be seen as the average performance of all the policies tried by the algorithm. This is particularly meaningful in the context of online learning that we are considering in this paper.

>>>>

%Paper structure
After providing background on policy gradient and SVRG in Section \ref{sec:pre}, in Section \ref{sec:alg} we propose SVRPG, a variant of SVRG for the policy gradient framework, addressing all the difficulties mentioned above. In Section \ref{sec:conv} we provide convergence guarantees for our algorithm, and we show a convergence rate that has an $O(\nicefrac{1}{T})$ dependence on the number $T$ of iterations. In Section \ref{sec:stopping} we suggest how to set the meta-parameters of SVRPG, while in Section \ref{sec:prac} we discuss some practical variants of the algorithm. Finally, in Section \ref{sec:exp} we empirically evaluate the performance of our method on popular continuous RL tasks.

@@

\vspace{-0.05in}
\subsection{Stochastic Variance-Reduced Gradient}\label{sec:svrg}
\vspace{-0.05in}
Finite-sum optimization is the problem of maximizing an objective function $f(\vtheta)$ which can be decomposed into the sum or average of a finite number of functions $z_i(\cdot|\vtheta)$:
\begin{align*}
        \max_{\vtheta} \left\{ f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}z_i(\vtheta)\right\}.
\end{align*}
This kind of optimization is very common in machine learning, where each $z_i$ may correspond to a data sample $x_i$ from a dataset $\mathcal{D}_N$ of size $N$ (\ie $z_i(\vtheta) = z(x_i|\vtheta)$). 
%% In this case, we adopt the following, more meaningful notation:
%% \begin{align*}
%% \max_{\vtheta} f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}g(x_i \vert \vtheta).
%% \end{align*}
A common requirement is that $z$ must be smooth and concave in $\vtheta$.\footnote{Note that we are considering a maximization problem instead of the classical minimization one.} 
Under this hypothesis, full gradient (FG) ascent~\citep{cauchy1847methode} with a constant step size achieves a linear convergence rate in the number $T$ of iterations (\ie parameter updates)~\citep{nesterov2013introductory}.
However, each iteration requires $N$ gradient computations, which can be too expensive for large values of $N$. Stochastic Gradient (SG) ascent~\citep[\eg][]{robbins1951stochastic,bottou2004large} overcomes this problem by sampling a single sample $x_i$ per iteration, but a vanishing step size is required to control the variance introduced by sampling. As a consequence, the lower per-iteration cost is paid with a worse, sub-linear convergence rate \cite{nemirovskii1983problem}.
Starting from SAG, a series of variations to SG have been proposed to achieve a better trade-off between convergence speed and cost per iteration: \eg SAG~\citep{roux2012stochastic}, SVRG~\cite{johnson2013accelerating}, SAGA~\cite{defazio2014saga}, Finito~\cite{defazio2014finito}, and MISO~\cite{mairal2015incremental}. 
The common idea is to reuse past gradient computations to reduce the variance of the current estimate.
In particular, Stochastic Variance-Reduced Gradient (SVRG) is often preferred to other similar methods for its limited storage requirements, which is a significant advantage when deep and/or wide neural networks are employed.

@@

\vspace{-0.05in}
\section{SVRG in Reinforcement Learning}\label{sec:alg}
\vspace{-0.05in}
In online RL problems, the usual approach is to tune the batch size of SG to find the optimal trade-off between variance and speed.
Recall that, compared to SL, the samples are not fixed in advance but we need to collect them at each policy change.
Since this operation may be costly, we would like to minimize the number of interactions with the environment.
For these reasons, we would like to apply SVRG to RL problems in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence.
However, a direct application of SVRG to RL is not possible due to the following issues:
% In on-line policy gradient problems, SG is not really a choice, but is dictated by the necessity of interacting with an unknown environment.\todopir{I do not understand the meaning of this sentence. I would say that SG is not a choice by referring to papers that use batches}
% We would like to apply SVRG to the policy gradient framework in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence. In this scenario, the function $f(\vtheta)$ to optimize is $J(\vtheta)$ and the (implicit) dataset $\mathcal{D}$ is the set of all possible trajectories $\mathcal{T}$. Compared to the typical finite sum optimization scenario, we have the following additional challenges:
\begin{description}
        \item[Non-concavity:] the objective function $J(\vtheta)$ is typically non-concave.
        \item[Infinite dataset:] the RL optimization cannot be expressed as a finite-sum problem. The objective function is an expected value over the trajectory density $p_{\vtheta}(\tau)$ of the total discounted reward, for which we would need an infinite dataset.
%    \item[Approximation:] the dataset $\mathcal{T}$ can be infinite, but we can only sample a finite number of trajectories;
        \item[Non-stationarity:] the distribution of the samples changes over time. In particular, the value of the policy parameter $\vtheta$ influences the sampling process.
\end{description}
To deal with non-concavity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as Gaussian\footnote{See Appendix~\ref{app:gauss} for more details on the Gaussian policy case.} and softmax~\citep[\eg][]{Furmston2012unifying,pirotta2015lipschitz}.
Because of the infinite dataset, we can only rely on an estimate of the full gradient.
\citet{harikandeh2015stopwasting} analysed this scenario under the assumptions of $z$ being concave, showing that SVRG is robust to an inexact computation of the full gradient. In particular, it is still possible to recover the original convergence rate if the error decreases at an appropriate rate. \citet{bietti2017stochastic} performed a similar analysis on MISO. In Section \ref{sec:conv} we will show how the estimation accuracy impacts on the convergence results with a non-concave objective.
Finally, the non-stationarity of the optimization problem introduces a bias into the SVRG estimator in Eq.~\eqref{E:svrg.gradient.correction}.
To overcome this limitation we employ importance weighting~\citep[\eg][]{rubinstein1981simulation,precup2000eligibility} to correct the distribution shift.
%Finally, we employ importance weighting~\citep[\eg]{rubinstein1981simulation,precup2000eligibility} to guarantee proper sampling of trajectories.

\begin{algorithm}[tb]
	\caption{SVRPG}
	\label{alg:svrpg}
	\begin{algorithmic}
		\STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, batch size $N$, mini-batch size $B$, gradient estimator $g$, initial parameter $\vtheta_{m}^0 \coloneqq \wt{\vtheta}^0 \coloneqq \vtheta_0$
		\FOR{$s=0$ {\bfseries to} $S-1$}
		\STATE $\vtheta_0^{s+1} := \wt{\vtheta}^{s} = \vtheta_{m}^s$
		\STATE Sample $N$ trajectories $\{\tau_j\}$ from $p(\cdot\vert\wt{\vtheta}^{s})$
		\STATE $ \wt{\mu} = \gradApp{\wt{\vtheta}^{s}}{N}$ (see Eq.~\eqref{E:policygradient.estimate})% = \frac{1}{N}\sum_{j=0}^{N-1}g(\tau_j | \wt{\vtheta}^s)$
		\FOR{$t=0$ {\bfseries to} $m-1$}
		\STATE Sample $B$ trajectories $\{\tau_i\}$ from $p(\cdot\vert\vtheta_t^{s+1})$
		%        \STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1} \left( g(\tau_i|\vtheta_t^{s+1}) - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \right)$
		\STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1}
		\begin{aligned}[t]
		\Big( & g(\tau_i|\vtheta_t^{s+1})\\ 
		& \quad{} - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \Big)
		\end{aligned}$
		%        \STATE $\begin{aligned}[t]
		%                &v^{s+1}_t = \wt{\mu}\\ 
		%                & +\frac{1}{B} \sum\limits_{i=0}^{B-1} \left( g(\tau_i|\vtheta_t^{s+1}) - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \right)
		%                \end{aligned}$
		\STATE $v^{s+1}_t = \wt{\mu} + c^{s+1}_t$ % \gradApp{\wt{\vtheta}^{s}}{N} + c^{s}_t$
		%%		\begin{align*}
		%%		\blacktriangledown J(\vtheta_t^{s+1}) = 
		%%		&\gradApp{\wt{\vtheta}^s}{N} \\
		%%		&+\frac{1}{B}\sum_{i=0}^{B-1}\left[ 
		%%		\score{}{\tau_i\vert\vtheta_t^{s+1}}\Reward(\tau_i)\right. \\
		%%		&\left. - \omega(\tau_i)\score{}{\tau_i \vert \wt{\vtheta}^{s}}\Reward(\tau_i)\right]
		%%		\end{align*}
		\STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v^{s+1}_t$
		\ENDFOR
		\ENDFOR
		\STATE {\bfseries return} $\vtheta_A\coloneqq\vtheta_t^{s+1}$ with $(s,t)$ picked uniformly at random from $\{[0,S-1]\times[0,m-1]\}$
	\end{algorithmic}
\end{algorithm}

@@

\textit{3) Non-stationarity.} 
Similarly to what is done in SL~\citep{cortes2010learning}, we require the variance of the importance weight to be bounded.
	\begin{restatable}[On the variance of importance weights]{assumption}{varweights}\label{ass:M2}
		There is a constant $W<\infty$ such that, for each pair of policies encountered in Algorithm~\ref{alg:svrpg} and for each trajectory,
		\[
                \mathbb{V}ar\left[\omega(\tau| \vtheta_1, \vtheta_2)\right] \leq \VARIS, \quad \forall \vtheta_1,\vtheta_2 \in \realspace^d , \tau \sim p(\cdot|\vtheta_1).
		\]
	\end{restatable}
Differently from Assumptions~\ref{ass:bounded_score} and ~\ref{ass:REINFORCE}, Assumption~\ref{ass:M2} must be enforced by a proper handling of the epoch size $m$.

We can now state the convergence guarantees for SVRPG.
\begin{restatable}[Convergence of the SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
Assume the REINFORCE or the G(PO)MDP gradient estimator is used in SVRPG (see Equation~\eqref{E:svrpg.estimate.batch}).
Under Assumptions~\ref{ass:bounded_score}, \ref{ass:REINFORCE} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Algorithm~\ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,\zeta, \xi$ and for proper choice of the step size $\alpha$ and the epoch size $m$, the following property:
\begin{align*}
	&\EVV[]
	{\norm[2]{\nabla J(\vtheta_A)}^2} 
		\leq
		\frac{J(\vtheta^*)-J(\vtheta_0)}{\psi T} +
		\frac{\zeta}{N}
		+\frac{\xi}{B},
\end{align*}
where $\vtheta^*$ is a global optimum and $\psi,\zeta,\xi$ depend only on $\GRADLOG,\HESSLOG,\VARRF,\VARIS,\alpha$ and $m$.
\end{restatable}
% Specific constants and meta-parameter constraints are provided in the proof of the theorem in Appendix \ref{app:proofs}.
Refer to Appendix~\ref{app:proofs} for a detailed proof involving the definition of the constants and the meta-parameter constraints.
%The theoretical settings may be too conservative.
By analysing the upper-bound in Theorem~\ref{theo:convergence} we observe that: I) the $\mathcal{O}(\nicefrac{1}{T})$ term is coherent with results on non-concave SVRG~\citep[\eg][]{reddi2016stochastic}; II) the $\mathcal{O}(\nicefrac{1}{N})$ term is due to the FG approximation and is analogous to the one in~\citep{harikandeh2015stopwasting}; III) the $\mathcal{O}(\nicefrac{1}{B})$ term is due to importance weighting.
To achieve asymptotic convergence, the batch size $N$ and the mini-batch size $B$ should increase over time. 
In practice, it is enough to choose $N$ and $B$ large enough to make the second and the third term negligible, \ie to mitigate the variance introduced by FG approximation and importance sampling, respectively.
Once the last two terms can be neglected, the number of trajectories needed to achieve $\norm[2]{\gradJ{\vtheta}}^2\leq\epsilon$ is $O(\frac{B+\nicefrac{N}{m}}{\epsilon})$. In this sense, an advantage over batch gradient ascent can be achieved with properly selected meta-parameters. In Section~\ref{sec:stopping} we propose a joint selection of step size $\alpha$ and epoch size $m$.
Finally, from the return statement of Algorithm \ref{alg:svrpg}, it is worth noting that $J(\vtheta_A)$ can be seen as the average performance of all the policies tried by the algorithm. This is particularly meaningful in the context of online learning that we are considering in this paper.
